---
title: "JSC370 Final Project"
author: "Runshi Yang"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>

This is my JSC370 Final Project website. I will showcase a few interactive visuals here.

<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(leaflet)
library(sf)
library(ggplot2)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```


```{r load-data, echo=FALSE}
KSI <- read.csv("https://raw.githubusercontent.com/Runshi-Yang/JSC370/main/assignments/a5/KSI.csv")
```


```{r plot1, class.source="code-r-small", echo=FALSE, results = 'hide'}
# Load neighborhood shapefile
neighborhoods <- st_read("Neighbourhoods.geojson")

# Filter KSI data for the time period and location covered by the neighborhood shapefile
KSI_filtered <- KSI %>%
  filter(YEAR >= 2006 & YEAR <= 2021) %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

# Aggregate KSI data by neighborhood
collisions_by_neighborhood <- st_join(KSI_filtered, neighborhoods, join = st_intersects) %>%
  group_by(AREA_NAME) %>%
  summarize(count = n()) %>%
  mutate(AREA_NAME = as.character(AREA_NAME))  # Add AREA_NAME column

# Merge neighborhood and collision data
neighborhoods_collisions <- st_join(neighborhoods, collisions_by_neighborhood)

# Create a leaflet map
pal <- colorNumeric(palette = "Reds", domain = neighborhoods_collisions$count)
plot1 <- leaflet(neighborhoods_collisions) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(fillColor = ~pal(count),
              fillOpacity = 0.8,
              color = "white",
              weight = 1,
              popup = ~paste(AREA_NAME.x, "<br>", "Number of Collisions: ", count)) %>%
  addLegend(position = "bottomright", 
            title = "Number of Collisions",
            pal = pal,
            values = neighborhoods_collisions$count)
```
```{r plot2, class.source="code-r-small", echo=FALSE}
KSI_filtered1 <- KSI %>%
  select(YEAR, ROAD_CLASS) %>%
  group_by(YEAR, ROAD_CLASS) %>%
  summarise(count = n())

plt_ksi <- ggplot(KSI_filtered1, aes(x = YEAR, y = count, color = ROAD_CLASS)) +
  geom_line() +
  labs(title = "Number of KSI Collisions by Road Type and Year",
       x = "Year",
       y = "Number of Collisions",
       color = "Road Type") +
  theme_minimal()

plot2 <- ggplotly(plt_ksi)
```


```{r plot3, class.source="code-r-small", echo=FALSE}
# Filter KSI data for alcohol-related collisions only
KSI_alcohol <- KSI %>%
  filter(ALCOHOL == TRUE)

# Group the data by year and injury level, and calculate the count of alcohol-related collisions
KSI_alcohol_by_year <- KSI_alcohol %>%
  group_by(YEAR, INJURY) %>%
  summarize(count = n())

# Create a ggplot object with year on the x-axis, count on the y-axis, and injury level as color
plt_alcohol <- ggplot(KSI_alcohol_by_year, aes(x = YEAR, y = count, color = INJURY)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Alcohol-Related Collisions by Year and Injury Level", 
       x = "Year", y = "Count", color = "Injury Level")

# Convert the ggplot object to a plotly object
plot3 <- ggplotly(plt_alcohol)
```
## Interactive Visulations {.tabset}

Create tabs to display each figure

### Map of Toronto KSI

```{r p1, echo=FALSE}
plot1
```
The map of Toronto KSI shows that West Humber-Clairville is the most crash-prone neighborhood, followed by the Yonge-Bay Corridor and South Riverdale in downtown. Wexford/Maryvale, Milliken and York University Heights has also experienced a relatively high number of traffic accidents.
<br>


### KSI by rode type for each year

```{r p2, echo=FALSE}
plot2
```
This plot shows that Major and Minor Arterial are the most frequent road types for traffic accidents, and the number of accidents on these two roads gradually decreases as the year progresses. The number of traffic accidents on other roads remained more or less the same.
<br>

### Alcohol related KSI
```{r p3, echo=FALSE}
plot3
```
This graph shows the trend of crashes related to drunk driving over the years. We can see a rapid decrease in the frequency of drunk driving crashes from 2006 onward, especially a significant decrease in cases that resulted in MAJOR injury and no INJURY.
<br>



